Water Management

### 2020-02-18 Paper/paper.tex
minor changes

In this work, we want to examine the variations in use of water from a (hypothetical) water reservoir.
One of the most important uses of surface water is the irrigation of fields for food production. However, as farming and food production are private enterprises and lakes are public goods, interests of all different parties have to be considered.
The water reservoir to be considered has two stakeholders: farmers, who want to use the water for irrigation purposes and citizens, who want to use the lake for recreational activities.
Note that while looking at just two agents is a vast simplification compared to a real allocation problem, one can easily add more stakeholders to the model, as long as there is enough computational power.
The sequence of events for every period is displayed in ~\ref{fig:Sequence}.
At the beginning of every year (Step 1) we have to decide how much of the water from the reservoir is used for the irrigation of fields, which benefits the farmers' harvest.\\\\
The farmers utility function follows the form

@@

with the additional paramter $s$ that indicates the water level at the beginning of the year.
For recreational users, the preference parameters shell be $a_2 = 2$ and $b_2 = -3$.\\\\

@@

where the weights $w_i$ are given as the solution to a linear system of equations and the nodes $x_i$ correspond to the zeros of the nth Hermite polynomial. (\cite{seminar:week6})

@@

In a first step, we want to compute the value function $V$ for each state $s$. 
In order to do so, we need to calculate an auxiliary matrix $aux$ that contains the combined utilities of the farmers and the recreational users. 
Obviously, the $aux$ matrix can only a finite, discrete amount of values. In this work, a resolution of $n = 1000$ different water levels and irrigation amounts should suffice. 
Thus we discretize both water level $s$ and irrigation amount $x$ as vectors

@@

Note that the maximum reservoir capacity is $7$ and $|s|=|x| = n = 1000$.

@@

\begin{equation}
V(s) = \max_{x}(u(s,x) + \beta*V'(s')))
\end{equation}

which equals to finding the column $aux$ that maximizes the expected value over all states (TODO: kann man das so sagen?).
We repeat this step over and over again until the our value function $V$ converges (technically: the norm of the difference between $V'$ and $V$ falls below a certain tolerance value). \\

@@

The value function as posed in % TODO put links to theory here
contains the maximum possible usage that can be gained by irrigating a specific amount of water under a given water level. 

@@

Every value along the possible water levels leads to an irrigation amount that optimizes the value for all parties in the given situation. Summarizing these values over the range of possible water levels results in a optimal irrigation policy which is shown in \ref{fig:optimal-irrigation-policy}.

@@

Due to the fact, that our water level highly depends on the randomly distributed rainfalls that fill the reservoir at end of each period, it is not possible to find a state steady. % TODO: auch noch evaporation nennen?
To accomplish this, we defined our steady state, to be a state, where the mean water level of the last 200 periods did not change more than $1x10^{-7}$ from a period to the next.

>>>>

In this work, we want to examine the variations in use of water from a (hypothetical) water reservoir.
One of the most important uses of surface water is the irrigation of fields for food production. However, as farming and food production are private enterprises and lakes are public goods, interests of all different parties have to be considered.
The water reservoir to be considered has two stakeholders: farmers, who want to use the water for irrigation purposes and citizens, who want to use the lake for recreational activities.
Note that while looking at just two agents is a vast simplification compared to a real allocation problem, one can easily add more stakeholders to the model, as long as there is enough computational power.
The sequence of events for every period is displayed in \ref{fig:Sequence}.
At the beginning of every year (Step 1) we have to decide how much of the water from the reservoir is used for the irrigation of fields, which benefits the farmers' harvest.\\\\
The farmers utility function follows the form

@@

with the additional paramter $s$ that indicates the water level at the beginning of the year.
For recreational users, the preference parameters shall be $a_2 = 2$ and $b_2 = -3$.\\\\

@@

where the weights $w_i$ are given as the solution to a linear system of equations and the nodes $x_i$ correspond to the zeros of the n-th Hermite polynomial. (\cite{seminar:week6})

@@

In a first step, we want to compute the value function $V$ for each state $s$. 
In order to do so, we need to calculate an auxiliary matrix $aux$ that contains the combined utilities of the farmers and the recreational users. 
Obviously, the $aux$ matrix can only contain a finite, discrete amount of values. In this work, a resolution of $n = 1000$ different water levels and irrigation amounts should suffice. 
Thus we discretize both water level $s$ and irrigation amount $x$ as vectors

@@

% TODO das scheint mir nicht so ganz korrekt zu sein. Meint ihr die Dimension?
Note that the maximum reservoir capacity is $7$ and $|s|=|x| = n = 1000$.

@@

\begin{equation}
V(s) = \max_{x}(u(s,x) + \beta*V'(s')))
\end{equation}
\label{eq:value-function}

which equals to finding the column $aux$ that maximizes the expected value over all states (TODO: kann man das so sagen?).
We repeat this step over and over again until the our value function $V$ converges.
In our case the function converges, if the norm of the difference between $V'$ and $V$ falls below a certain tolerance value) \\

@@

The value function as posed in \ref{eq:value-function} contains the maximum possible usage that can be gained by irrigating a specific amount of water under a given water level. 

@@

Every value along the possible water levels is mapped to an irrigation amount that optimizes the value for all parties in the given situation. Summarizing these values over the range of possible water levels results in a optimal irrigation policy which is shown in \ref{fig:optimal-irrigation-policy}.

@@

Due to the fact, that our water level highly depends on the randomly distributed rainfalls that fill the reservoir at end of each period, it is not possible to find a steady state in which the water level does not change at all. % TODO: auch noch evaporation nennen?
To analyze it nevertheless, we defined our steady state, to be a state, where the mean water level of the last 200 periods did not change more than $1x10^{-7}$ from a period to the next.
